%  ========================================================================
%  Copyright (c) 1985 The University of Washington
%
%  Licensed under the Apache License, Version 2.0 (the "License");
%  you may not use this file except in compliance with the License.
%  You may obtain a copy of the License at
%
%      http://www.apache.org/licenses/LICENSE-2.0
%
%  Unless required by applicable law or agreed to in writing, software
%  distributed under the License is distributed on an "AS IS" BASIS,
%  WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
%  See the License for the specific language governing permissions and
%  limitations under the License.
%  ========================================================================
%

% Documentation for University of Washington thesis LaTeX document class
% by Jim Fox
% fox@washington.edu
%
%    Revised 2020/02/24, added \caption()[]{} option.  No ToC.
%
%    Revised for version 2015/03/03 of uwthesis.cls
%    Revised, 2016/11/22, for cleanup of sample copyright and title pages
%
%    This document is contained in a single file ONLY because
%    I wanted to be able to distribute it easily.  A real thesis ought
%    to be contained on many files (e.g., one for each chapter, at least).
%
%    To help you identify the files and sections in this large file
%    I use the string '==========' to identify new files.
%
%    To help you ignore the unusual things I do with this sample document
%    I try to use the notation
%       
%    % --- sample stuff only -----
%    special stuff for my document, but you don't need it in your thesis
%    % --- end-of-sample-stuff ---


%    Printed in twoside style now that that's allowed
%
 
\documentclass[11pt, proquest]{uwthesis}[2016/11/22]

\usepackage[hidelinks]{hyperref}
\usepackage[round]{natbib}
\usepackage{graphicx}
\usepackage{notoccite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{tabularx}
\usepackage{ltablex}
\usepackage{setspace}
\usepackage{bm}
\usepackage{soul}

% for CV
\usepackage{enumitem}
\newlist{cvlist}{description}{1}
\setlist[cvlist]{style=sameline,leftmargin=1in,font=\normalfont}

\renewcommand{\arraystretch}{1.0}

%
% The following line would print the thesis in a postscript font 

% \usepackage[round]{natbib}
% \def\bibpreamble{\protect\addcontentsline{toc}{chapter}{Bibliography}}

\setcounter{tocdepth}{2}  % Print the chapter and sections to the toc
 
% ==========   Local defs and mods
%

\begin{document}
 
% ==========   Preliminary pages
%
% ( revised 2012 for electronic submission )
%

\prelimpages
 
%
% ----- copyright and title pages
%
\Title{Biosignatures, the Origin of Life, and the Early Earth Atmosphere}
\Author{Nicholas Wogan}
\Year{2023}
\Program{Earth and Space Sciences}
\Chair{David C. Catling}{Professor}{Earth and Space Sciences}
\Signature{Kevin J. Zahnle}
\Signature{Roger Buick}
\Signature{Rory Barnes}

\copyrightpage

\titlepage  

 
%
% ----- signature and quote are gone
%

%
% ----- abstract
%


\setcounter{page}{-1}
\include{tex/0abstract.tex}
 
%
% ----- contents & etc.
%
\tableofcontents
\listoffigures
\listoftables
 
%
% ----- acknowledgments
%
\acknowledgments{

This dissertation was made possible by a long list of mentors, graduate students, staff, friends and family.

% Committee
First I would like to thank my advisor, David Catling. I have been very fortunate to benefit from David's vast knowledge of astrobiology and planetary science. I could not have asked for a better advisor. I also need to thank my mentor and committee member Kevin Zahnle. Almost all of my dissertation chapters were improved by Kevin's intuition for how planet work. Additionally, I thank the rest of my committee members, Rory Barnes and Roger Buick, for providing me with constructive feedback on my research over the last few years.

% Josh KT
This thesis would not have been successful without Josh Krissansen-Totton. Thanks for being a great friend, collaborator, and mentor.

% Astrobiology program and Vikki
The Astrobiology Program has been an integral part of my experience at the University of Washington. Thanks to the director Victoria Meadows, and the many staff and students that kept the program thriving while I was a part of it.

% List of other mentors and collaborators
I also need to thank the following friends, colleague or mentors for helping me at various stages of my PhD: Andrew Hoffman, Gemma O`Conner, Lucas Fifer, Trent Thomas, Zoe Todd, Zack Cohen, Ping-Chun Lin, Andrew Shumway, Maggie Thompson, Shintaro Kadoya, Eric Wolf, Shawn Domagal-Goldman, Sandra Bastelberger, Chris Stark, Dale Winebrenner, Jake Lustig-Yaeger, and Andrew Lincowski. 

% Thanks to family and Emily.
My family has always given me unconditional love and support. Thank you Mom, Dad and Ted. Finally, I need to thank Emily Honea for consistently being there for me.

}

%
% ----- dedication
%
\dedication{\begin{center}
  To my parents, Cam and Mary Lou, and my brother, Ted.
\end{center}}

%
% end of the preliminary pages
 
%
% ==========      Text pages
%

\textpages

% ========== Chapters

% Chapter 1
\chapter{Introduction}
\noindent \emph{``It has been said that astronomy is a humbling and character-building experience. There is perhaps no better demonstration of the folly of human conceits than this distant image of our tiny world.''} - Carl Sagan, in the Pale Blue Dot
\newpage
\include{tex/1intro.tex}

% Part 1
\part{Atmospheric Biosignatures, and Life's Impact on the Early Earth} \label{pt:1}
 
% Chapter 2
\chapter{Atmospheric Chemical Disequilibrium from Dead to Living Worlds} \label{ch:2}
\noindent \emph{``It may be some time before we can try these tests for life on the planets of other stars but in the meanwhile what started as an exercise in exobiology could become an expedition to discover the largest living creature on Earth, Gaia.''} - \citet{Lovelock_1975}
\newpage
\include{tex/2diseq.tex}

% Chapter 3
\chapter{The Likelihood of CH\textsubscript{4}+CO\textsubscript{2} Biosignature False Positives from Volcanism} \label{ch:3}
\noindent \emph{``Life is the hypothesis of last resort.''} - \citet{Sagan_1993}
\newpage
\include{tex/3methane.tex}

% Chapter 4
\chapter{Rapid timescale for an oxic transition during the Great Oxidation Event and the instability of low atmospheric O\textsubscript{2}} \label{ch:4}
\noindent \emph{``To evaluate $f_l$, the fraction of such planets on which life develops, we perform a simple series of spectrographic observations. We first search for large amounts of molecular oxygen - a sure sign that something strange is happening on that planet.''} - \citet{Owen_1980}

\newpage
\include{tex/4goe.tex}

% Part 2
\part{Origin of life atmospheric chemistry} \label{pt:2}

% Chapter 5
\chapter{Origin of Life Molecules in the Atmosphere After Big Impacts on the Early Earth} \label{ch:5}
\noindent \emph{``At the time such objects were falling on the moon similar objects fell on the Earth... Its materials would have fallen through the atmosphere in the form of iron and silicate rains and would have reacted with the atmosphere in the process... Thus it is very difficult to see how the primitive atmosphere of the Earth contained more than trace amounts of other compounds of carbon, nitrogen, oxygen and hydrogen than CH$_4$, H$_2$O, NH$_3$ (or N$_2$) and H$_2$.''} - \citet{Urey_1952}
\newpage
\include{tex/5impacts.tex}

% Chapter 6
\chapter{Timing and Likelihood of the Origin of Life Derived from Post-Impact Highly Reducing Atmospheres} \label{ch:6}
\noindent \emph{``At some point a particularly remarkable molecule was formed by accident. We will call it the Replicator. It may not necessarily have been the biggest or the most complex molecule around, but it had the extraordinary property of being able to create copies of itself. This may seem a very unlikely sort of accident to happen. So it was. It was exceedingly improbable... But in our human estimates of what is probable and what is not, we are not used to dealing in hundreds of millions of years.''} - Richard Dawkins in The Selfish Gene
\newpage
\include{tex/6time.tex}

% Chapter 7
\chapter{Conclusions}
\noindent \emph{``The purpose of a storyteller is not to tell you how to think, but to give you questions to think upon.''} - Brandon Sanderson
\newpage
\include{tex/7conclusions.tex}

\printendnotes

%
% ==========   Bibliography
%
% \nocite{*}
\bibliographystyle{plainnat}
\bibliography{bib}
%
% ==========   Appendices
%
\appendix
\raggedbottom\sloppy
 
% ========== Appendix A
 
\chapter{Photochemical Modeling Equations}
\newpage
\include{tex/A_appendix.tex}

\chapter{Curriculum Vitae}
\newpage
\include{tex/cv.tex}
 
\end{document}